---
title: "Model Building"
author: "Jeremy Chan"
date: "June 13, 2016"
output: html_document
---

The aim of this document is to catalog the attempts at creating models to predict if passengers on the Titanic survived. The predicton will be based on a training set and then evaluated on a test set for a Kaggle competition.

Loading data
```{r}
if(Sys.info()["sysname"]=="Darwin")
{setwd("~/Documents/GitHub/Kaggle-Titanic/Data")}
# add windows working directory

train = read.csv("train.csv")
test = read.csv("test.csv")
```

Data exploration
```{r}
str(train)
```
891 observations of 12 variables
  PassengerID, survived, passenger class, name, sex, age, num of sibling/spouses aboard, number of parents/children aboard, ticket number, passenger fare, cabin, port of embarkation (C = Cherbourg, Q = Queenstown, S = Southampton)
  
Survival Rates
```{r}
table(train$Survived, dnn = "Survived")
prop.table(table(train$Survived, dnn = "Survived"))
```

It looks like there is a 38.38% survival rate among all passengers. 

Set up test survival rates and create ouput file 
```{r}
# assume everyone died
test$Survived <- rep(0, 418)
submit = data.frame(PassengerId = test$PassengerId, Survived = test$Survived)
write.csv(submit, file = "theyalldie.csv", row.names = FALSE)
```

Adding more varibles

Exploring gender

Gender Survival Rates
```{r}
table(train$Sex)
prop.table(table(train$Survived, train$Sex), 1)
```

68.13% of females survived and 31.87% of males survived. 